Chemistry of the Orphan Stream I: Identifying Stream Candidates with Low Resolution Spectroscopy

### 2012-09-12 2012-oss-mrs/draft-oss-mrs.tex
Current status of paper prior to applying GDC comments

\author{Andrew R. Casey\altaffilmark{1,2}, Stefan C. Keller\altaffilmark{1}, Gary Da Costa\altaffilmark{1}, Elizabeth Maunder\altaffilmark{1}, Anna Frebel\altaffilmark{2}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; \email{arcasey@mso.anu.edu.au}}
\altaffiltext{2}{Massachusetts Institute of Technology, Kavli Institute for Astrophysics and Space Research,
77 Massachusetts Avenue, Cambridge, MA 02139, USA}

@@

The Orphan Stream was independently detected by both \citet{Grillmair;Dionatos_2006} and \citet{Belokurov;et-al_2006}, and is distinct from other substructures in the halo. The stream stretches over $60\,^\circ$ in the sky, has a low surface brightness and a narrow stream width of only $\sim2^\circ$ and as the name suggests, the parent object largely remains a mystery. The stream extends past the celestial equator \-- outside the SDSS footprint \-- but is not detected in existing southern surveys \citep{Newberg;et-al_2010}. Whilst the parent system remains elusive, significant effort has been placed on associating the stream with known Milky Way satellites \citep{Zucker;et-al_2006, Fellhaur;et-al_2007,Jin;Lynden_Bell_2007,Sales;et-al_2008}. In contrast, there has been relatively limited observational work on the Orphan Stream itself other than the original discovery papers \citep{Grillmair;Dionatos_2006, Belokurov;et-al_2006, Belokurov;et-al_2007} and the work of \citet{Newberg;et-al_2010}. This is largely to be expected given the absence of multi-band photometry in the southern sky and the total low luminosity of the stream, making it difficult in reliably identifying Orphan Stream members from halo stars. It is unlikely the full extent of the Orphan Stream will be detected until the SkyMapper telescope surveys this region \citep{Keller;et-al_2007}.

@@

% Besan\c{c}on heliocentric velocities have been translated to a galactocentric frame in the same way as our targets.
Figure \ref{fig:velocities} shows a histogram of our galactocentric velocities, compared to the predicted smooth halo velocity distribution for this region from the Besan\c{c}on model \citep{Robin;et-al_2003}. It is clear that our target selection has yielded mostly nearby thick disk dwarf stars with $V_{\textsc{gsr}} \approx -120$\,km s$^{-1}$. The expected kinematic signature for the Orphan Stream, as \citet{Newberg;et-al_2010} reported, is labelled in Figure \ref{fig:velocities}. No statistically significant peak representative of the Orphan Stream exists in our sample. From kinematics alone, our targets appears largely indistinguishable from a smooth halo distribution.

@@

This analysis was also performed using the total equivalent width of the Mg\,\textsc{I}\,b triplet lines. Both analyses were entirely consistent with each other: essentially the same candidate selection was found using both techniques. However, given slightly poorer signal at the Mg\,\textsc{I}\,b triplet, we were forced to adopt many more upper limits than when using the 8807{\AA} line. Because we classify all upper limits as being ``non-dwarfs" (i.e. potential OSS giants), we deduced a slightly larger candidate sample for the Mg\,\textsc{I}\,b analysis, which was primarily populated by upper limits. In conclusion, we found the Mg\,\textsc{I} line at 8807 appeared to be a more consistent dwarf discriminant given our weak S/N \--- particularly for our fainter stars \--- and thus we have used the 8807{\AA} Mg \textsc{I} selection throughout the rest of our analysis.

@@

We can ascertain abundance corrections for our stars fainter than the horizontal branch by comparing our $V-V_{HB}$ and $EW_{\lambda8452} + EW_{\lambda8662}$ with those measured by \citet{Saviane;et-al_2012} for globular clusters. Only two clusters had measured Ca \textsc{II} equivalent widths for stars fainter than the horizontal branch: NGC 6397 with $[\mbox{Fe/H}] = -1.99$ and NGC 6254 with $[\mbox{Fe/H}] = -1.57$ \citep[both abundances from][]{Carretta;et-al_2009}. OSS 6 and OSS 7 lie at the mid-point between these cluster stars, implying an abundance of $[\mbox{Fe/H}] \sim -1.78$, or a mean correction of $-0.14$ dex for OSS 6 and 7. Although OSS 3 is 0.15 magnitudes fainter than the horizontal branch, it has equivalent widths consistent with our inferred metallicity of $[\mbox{Fe/H}] = -1.62$, so no correction was made. With corrections included for OSS 6 and 7 we find a median stream metallicity from our four most probable stars of $[\mbox{Fe/H}] = -1.78 \pm 0.08$, which is slightly more metal-rich than the $[\mbox{Fe/H}] =-2.1$ found by \citet{Newberg;et-al_2010} from BHB stars. If we include our EMP candidate, the median decreases by 0.01 dex and the spread increases to 0.39\,dex.

>>>>

\author{Andrew R. Casey\altaffilmark{1,2}, Stefan C. Keller\altaffilmark{1}, Gary Da Costa\altaffilmark{1}, Elizabeth Maunder\altaffilmark{1}, Anna Frebel\altaffilmark{2}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; \email{acasey@mso.anu.edu.au}}
\altaffiltext{2}{Massachusetts Institute of Technology, Kavli Institute for Astrophysics and Space Research,
77 Massachusetts Avenue, Cambridge, MA 02139, USA}

@@

The Orphan Stream was independently detected by both \citet{Grillmair;Dionatos_2006} and \citet{Belokurov;et-al_2006}, and is distinct from other substructures in the halo. The stream stretches over $60\,^\circ$ in the sky, has a low surface brightness and a narrow stream width of only $\sim2^\circ$ and as the name suggests, the parent object largely remains a mystery. The stream extends past the celestial equator \--- outside the SDSS footprint \--- but is not detected in existing southern surveys \citep{Newberg;et-al_2010}. Whilst the parent system remains elusive, significant effort has been placed on associating the stream with known Milky Way satellites \citep{Zucker;et-al_2006, Fellhaur;et-al_2007,Jin;Lynden_Bell_2007,Sales;et-al_2008}. In contrast, there has been relatively limited observational work on the Orphan Stream itself other than the original discovery papers \citep{Grillmair;Dionatos_2006, Belokurov;et-al_2006, Belokurov;et-al_2007} and the work of \citet{Newberg;et-al_2010}. This is largely to be expected given the absence of multi-band photometry in the southern sky and the total low luminosity of the stream, making it difficult in reliably identifying Orphan Stream members from halo stars. It is unlikely the full extent of the Orphan Stream will be detected until the SkyMapper telescope surveys this region \citep{Keller;et-al_2007}.

@@

% Besan\c{c}on heliocentric velocities have been translated to a galactocentric frame in the same way as our targets.
Figure \ref{fig:velocities} shows a histogram of our galactocentric velocities, compared to the predicted smooth line-of-sight velocity distribution for this region from the Besan\c{c}on model \citep{Robin;et-al_2003}. It is clear that our target selection has yielded mostly nearby thick disk dwarf stars with $V_{\textsc{gsr}} \approx -120$\,km s$^{-1}$. The expected kinematic signature for the Orphan Stream, as \citet{Newberg;et-al_2010} reported, is labelled in Figure \ref{fig:velocities}. No statistically significant peak representative of the Orphan Stream exists in our sample. From kinematics alone, our targets appears largely indistinguishable from a smooth halo distribution.

@@

This analysis was also performed using the total equivalent width of the Mg\,\textsc{I}\,b triplet lines. Both analyses were entirely consistent with each other: essentially the same candidate selection was found using both techniques. However, given slightly poorer signal at the Mg\,\textsc{I}\,b triplet, we were forced to adopt many more upper limits than when using the 8807 {\AA} line. Because we classify all upper limits as being ``non-dwarfs" (i.e. potential OSS giants), we deduced a slightly larger candidate sample for the Mg\,\textsc{I}\,b analysis, which was primarily populated by upper limits. In conclusion, we found the Mg\,\textsc{I} line at 8807 appeared to be a more consistent dwarf discriminant given our weak S/N \--- particularly for our fainter stars \--- and thus we have used the 8807{\AA} Mg \textsc{I} selection throughout the rest of our analysis.

@@

We can ascertain abundance corrections for our stars fainter than the horizontal branch by comparing our $g - g_{HB}$ (or $V-V_{HB}$) and $EW_{\lambda8452} + EW_{\lambda8662}$ with those measured by \citet{Saviane;et-al_2012} for globular clusters. Only two clusters from \citet{Saviane;et-al_2012} had measured Ca \textsc{II} equivalent widths for stars fainter than the horizontal branch: NGC 6397 with $[\mbox{Fe/H}] = -1.99$ and NGC 6254 with $[\mbox{Fe/H}] = -1.57$ \citep[both abundances from][]{Carretta;et-al_2009}. OSS 6 and OSS 7 lie at the mid-point between these cluster stars, implying an abundance of $[\mbox{Fe/H}] \sim -1.78$, or a mean correction of $-0.14$ dex for OSS 6 and 7. Although OSS 3 is 0.15 magnitudes fainter than the horizontal branch, it has equivalent widths consistent with our inferred metallicity of $[\mbox{Fe/H}] = -1.62$, so no correction was made. With corrections included for OSS 6 and 7 we find a median stream metallicity from our four most probable stars of $[\mbox{Fe/H}] = -1.78 \pm 0.08$, which is slightly more metal-rich than the $[\mbox{Fe/H}] =-2.1$ found by \citet{Newberg;et-al_2010} from BHB stars. If we include our EMP candidate, the median decreases by 0.01 dex and the spread increases to 0.39\,dex.
